---
title: "Creating Variables"
author: "Hugo van den Belt"
date: "2025-02-19"
output:
  pdf_document: default
  html_document: default
  word_document: default
---



# Creating Variables

## Load Libraries
```{r}
library(haven)
library(dplyr)
```

## Partisan Identifiers

Here I create dummies for identifying with a political party, based on the variable `partyid`.
This variable has 8 levels relating to how strongly someone identifies with a party. 
I make one variable where everyone who agrees slightly with either party (and is not independent) is labeled 
as that party, and one where only people who agree strongly are identified as that party.

The levels are: 
0. strong democrat
1. not very strong democrat
2. independent, close to democrat
3. independent (neither, no response)
4. independent, close to republican
5. not very strong republican
6. strong republican
7. other party


### Identifiers for any degree of identification

To create a variable where **any degree of party identification** is grouped under a party label, I define:  

- `Democrat_any = 1` if `partyid` is **0, 1, or 2** (Strong Democrat, Not Very Strong Democrat, Independent Close to Democrat).  
- `Republican_any = 1` if `partyid` is **4, 5, or 6** (Independent Close to Republican, Not Very Strong Republican, Strong Republican).  
- All others (pure independents, third-party voters) are coded as `0`.  


```{r}
subset_data_imputed <- subset_data_imputed %>%
  mutate(
    Democrat_any = ifelse(partyid %in% c(0, 1, 2), 1, 0),
    Republican_any = ifelse(partyid %in% c(4, 5, 6), 1, 0)
  )
```


### Identifiers for strong identification

To create a variable where **STRONG degree of party identification**:

```{r}
subset_data_imputed <- subset_data_imputed %>%
  mutate(
    Democrat_strong = ifelse(partyid == 0, 1, 0),
    Republican_strong = ifelse(partyid == 6, 1, 0)
  )
```


```{r}
table(subset_data_imputed$partyid)

table(subset_data_imputed$Democrat_any)
table(subset_data_imputed$Republican_any)

table(subset_data_imputed$Democrat_strong)
table(subset_data_imputed$Republican_strong)
```





## Variable Identifying the president's party

These are the years in the dataset

```{r}
unique(subset_data_imputed$year)
```

From the [White House Website], (https://www.whitehousehistory.org/the-presidents-timeline) we can see who was president in what year, leading to this table:

### **Party Affiliation of U.S. Presidents (1972–2022)**  

| Year  | President   | Party  |
|-------|------------|--------|
| 1972  | Nixon      | R      |
| 1973  | Nixon      | R      |
| 1974  | Ford       | R      |
| 1975  | Ford       | R      |
| 1976  | Ford       | R      |
| 1977  | Carter     | D      |
| 1978  | Carter     | D      |
| 1980  | Carter     | D      |
| 1982  | Reagan     | R      |
| 1983  | Reagan     | R      |
| 1984  | Reagan     | R      |
| 1985  | Reagan     | R      |
| 1986  | Reagan     | R      |
| 1987  | Reagan     | R      |
| 1988  | Reagan     | R      |
| 1989  | Bush Sr.   | R      |
| 1990  | Bush Sr.   | R      |
| 1991  | Bush Sr.   | R      |
| 1993  | Clinton    | D      |
| 1994  | Clinton    | D      |
| 1996  | Clinton    | D      |
| 1998  | Clinton    | D      |
| 2000  | Clinton    | D      |
| 2002  | Bush Jr.   | R      |
| 2004  | Bush Jr.   | R      |
| 2006  | Bush Jr.   | R      |
| 2008  | Bush Jr.   | R      |
| 2010  | Obama      | D      |
| 2012  | Obama      | D      |
| 2014  | Obama      | D      |
| 2016  | Obama      | D      |
| 2018  | Trump      | R      |
| 2021  | Biden      | D      |
| 2022  | Biden      | D      |

Now we can create a variable that is 1 when the president is Republican, and 0 if not (i.e., Democrat)

```{r}
# president's party dummy
subset_data_imputed <- subset_data_imputed %>%
  mutate(
    pres_republican = case_when(
      year %in% c(1972, 1973, 1974, 1975, 1976, 1982, 1983, 1984, 1985, 1986, 
                  1987, 1988, 1989, 1990, 1991, 2002, 2004, 2006, 2008, 2018) ~ 1,
      TRUE ~ 0
    )
  )

# Counts
table(subset_data_imputed$pres_republican)
```

```{r}
# Create a dummy for when the president is a Democrat (1 = Democrat, 0 = Republican)
subset_data_imputed <- subset_data_imputed %>%
  mutate(
    pres_democrat = case_when(
      year %in% c(1977, 1978, 1980, 1993, 1994, 1996, 1998, 2000, 
                  2010, 2012, 2014, 2016, 2021, 2022) ~ 1,
      TRUE ~ 0
    )
  )

# Check counts
table(subset_data_imputed$pres_democrat)
```


## Preparing (possible) outcome variables for ordered logistic regression
```{r}
library(dplyr)
library(MASS)

subset_data_imputed <- subset_data_imputed %>%
  mutate(finalter = factor(finalter, 
                           levels = c(2, 3, 1),  # Match numbers
                           labels = c("worse", "stayed same", "better"), 
                           ordered = TRUE),
         satfin = factor(satfin, 
                         levels = c(3, 2, 1),  
                         labels = c("not satisfied at all", "more or less satisfied", "pretty well satisfied"), 
                         ordered = TRUE))

# Check structure
levels(subset_data_imputed$finalter)
levels(subset_data_imputed$satfin)

str(subset_data_imputed$finalter)
str(subset_data_imputed$satfin)

```


## Increased Polarization Marker

```{r}
# Ensure 'year' is numeric
subset_data_imputed <- subset_data_imputed %>%
  mutate(year = as.numeric(as.character(year))) %>%  # Convert year from factor to numeric
  mutate(after_bushjr = if_else(year >= 2001, 1, 0))

# Check counts
table(subset_data_imputed$after_bushjr)
```


