---
title: "Importing Data Subset"
author: "Hugo van den Belt"
date: "2025-02-19"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

# Importing Data Subset

## Load Libraries
```{r}
library(haven)
library(dplyr)
```


## File path
```{r}
file_path <- "/Users/hugovandenbelt/Downloads/GSS_stata/gss7222_r4.dta"
gss_data <- read_dta(file_path)
```


The original Data Set can be downloaded from [the GSS website] (https://gss.norc.org/us/en/gss/get-the-data/stata.html) by downloading "Cumulative Data Set (cross-sectional samples from all years)"

## Select relevant variables
```{r}
subset_data <- gss_data %>%
  dplyr::select(year, id, marital, sibs, childs, age, educ, degree, padeg, 
         madeg, sex, race, reg16, income, rincome, region, partyid, polviews, 
         relig, happy, class, satfin, finalter, finrela)

subset_data <- subset_data %>%
  mutate(across(where(is.labelled), as.numeric))

# Convert necessary variables to factors
# It would be easier to subset directly as factors, but that messes up the NA observations, hence this workaround
factor_vars <- c("year", "marital", "educ", "degree", 
                 "padeg", "madeg", "sex", "race", "reg16", "income", "rincome", "region", 
                 "partyid", "polviews", "relig", "happy", "class", 
                 "satfin", "finalter", "finrela")

subset_data <- subset_data %>%
  mutate(across(all_of(factor_vars), as.factor))

head(subset_data)
```

## Check how many NAs exist in each column
```{r}
colSums(is.na(subset_data))
```

# Handling Missing Data

BEFORE RUNNING THE CHUNKS BELOW, ALL CHUNKS FROM THE VARIABLES MODULE SHOULD BE RUN 

There are some missing observations, but the package I'm using cannot run the logit model with missing observations. 

## Subset deleting missing observations

Deleting observations with missing values in final set of controls

```{r}
# Ensure missing values are correctly recognized
subset_data_clean <- subset_data %>%
  filter(complete.cases(finalter, age, degree, sex, race, income, class))

# check how many observations were removed
cat("Original dataset size:", nrow(subset_data), "\n")
cat("Cleaned dataset size:", nrow(subset_data_clean), "\n")
cat("Number of observations removed:", nrow(subset_data) - nrow(subset_data_clean), "\n")

head(subset_data_clean)

```


We must check if not too many observations have gone missing for a particular year

```{r}
table(subset_data$year)
table(subset_data_clean$year)
```

## FINAL dataset

Finally, we need to delete years where there is a transition of power

```{r}
# Remove observations from years 1977, 1993, and 2021
subset_final <- subset_data_clean %>%
  filter(!(year %in% c(1977, 1993, 2021)))

# Check how many observations were removed
cat("Original dataset size:", nrow(subset_data_clean), "\n")
cat("Final dataset size after removing years 1977, 1993, 2021:", nrow(subset_final), "\n")
cat("Number of observations removed:", nrow(subset_data_clean) - nrow(subset_final), "\n")

head(subset_final)

table(subset_final$year)

```

# Data subset post Bush
```{r}
# Remove all observations before 2001
subset_final_postbush <- subset_final %>%
  filter(year >= 2001)

# Check how many observations were removed
cat("Original dataset size:", nrow(subset_final), "\n")
cat("Final dataset size after removing years before 2001:", nrow(subset_final_postbush), "\n")
cat("Number of observations removed:", nrow(subset_final) - nrow(subset_final_postbush), "\n")

head(subset_final_postbush)

# Show the distribution of years in the cleaned dataset
table(subset_final_postbush$year)

```

## Save subsets as csv

```{r}
write.csv(subset_data, file = "/Users/hugovandenbelt/Documents/GitHub/Research Note EUR Recent Advancements/DataSets/subset_GSSdata.csv", row.names = FALSE)

write.csv(subset_data_clean, file = "/Users/hugovandenbelt/Documents/GitHub/Research Note EUR Recent Advancements/DataSets/subset_GSSdata_clean.csv", row.names = FALSE)

write.csv(subset_final, file = "/Users/hugovandenbelt/Documents/GitHub/Research Note EUR Recent Advancements/DataSets/subset_final.csv", row.names = FALSE)

write.csv(subset_final_postbush, file = "/Users/hugovandenbelt/Documents/GitHub/Research Note EUR Recent Advancements/DataSets/subset_final_postbush.csv", row.names = FALSE)
```


