---
title: "Importing Data Subset"
author: "Hugo van den Belt"
date: "2025-02-19"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

# Importing Data Subset

## Load Libraries
```{r}
library(haven)
library(dplyr)
```


## File path
```{r}
file_path <- "/Users/hugovandenbelt/Downloads/GSS_stata/gss7222_r4.dta"
gss_data <- read_dta(file_path)
```


The original Data Set can be downloaded from [the GSS website] (https://gss.norc.org/us/en/gss/get-the-data/stata.html) by downloading "Cumulative Data Set (cross-sectional samples from all years)"

## Select relevant variables
```{r}
subset_data <- gss_data %>%
  dplyr::select(year, id, marital, sibs, childs, age, educ, degree, padeg, 
         madeg, sex, race, reg16, income, rincome, region, partyid, polviews, 
         relig, happy, class, satfin, finalter, finrela)

subset_data <- subset_data %>%
  mutate(across(where(is.labelled), as.numeric))

# Convert necessary variables to factors
# It would be easier to subset directly as factors, but that messes up the NA observations, hence this workaround
factor_vars <- c("year", "marital", "educ", "degree", 
                 "padeg", "madeg", "sex", "race", "reg16", "income", "rincome", "region", 
                 "partyid", "polviews", "relig", "happy", "class", 
                 "satfin", "finalter", "finrela")

subset_data <- subset_data %>%
  mutate(across(all_of(factor_vars), as.factor))

head(subset_data)
```

## Check how many NAs exist in each column
```{r}
colSums(is.na(subset_data))
```

# Handling Missing Data

There are some missing observations, but the package I'm using cannot run the logit model with missing observations. 

## Subset deleting missing observations

Deleting observations with missing values in final set of controls

```{r}
# Ensure missing values are correctly recognized
subset_data_clean <- subset_data %>%
  filter(complete.cases(finalter, age, degree, sex, race, income, class))

# check how many observations were removed
cat("Original dataset size:", nrow(subset_data), "\n")
cat("Cleaned dataset size:", nrow(subset_data_clean), "\n")
cat("Number of observations removed:", nrow(subset_data) - nrow(subset_data_clean), "\n")

head(subset_data_clean)

```


We must check if not too many observations have gone missing for a particular year

```{r}
table(subset_data$year)
table(subset_data_clean$year)
```

## FINAL dataset

Finally, we need to delete years where there is a transition of power

```{r}
# Remove observations from years 1977, 1993, and 2021
subset_final <- subset_data_clean %>%
  filter(!(year %in% c(1977, 1993, 2021)))

# Check how many observations were removed
cat("Original dataset size:", nrow(subset_data_clean), "\n")
cat("Final dataset size after removing years 1977, 1993, 2021:", nrow(subset_final), "\n")
cat("Number of observations removed:", nrow(subset_data_clean) - nrow(subset_final), "\n")

head(subset_final)

table(subset_final$year)

```

# Data subset post Bush
```{r}
# Remove all observations before 2001
subset_final_postbush <- subset_final %>%
  filter(year >= 2001)

# Check how many observations were removed
cat("Original dataset size:", nrow(subset_final), "\n")
cat("Final dataset size after removing years before 2001:", nrow(subset_final_postbush), "\n")
cat("Number of observations removed:", nrow(subset_final) - nrow(subset_final_postbush), "\n")

head(subset_final_postbush)

# Show the distribution of years in the cleaned dataset
table(subset_final_postbush$year)

```



## Maximum Likelihood [ALTERNATIVE UNUSED APPROACH]

I ended up `not` using this method. Instead I just deleted observations with missing values (see beolow).

## Why Use Maximum Likelihood (ML) for Imputation?

Missing data is a problem when doing multinomial logistic regression. Instead of simply dropping observations with missing values, I use **Maximum Likelihood (ML) imputation** using **Multiple Imputation by Chained Equations (MICE)** for more reliable results.

I found most of the explanation on [this site] (https://www.blog.trainindata.com/multiple-imputation-with-chained-equations-mice-what-is-it/?)

### **Advantages of ML Imputation:**
1. **Preserves Data Relationships**  
   - Unlike mean or mode imputation ML imputation considers the relationships between variables when filling in missing values. This avoids distorting the structure of the data.

2. **Handles Both Categorical and Continuous Data**  
   - MICE allows different imputation methods for different variable types. This makes it more flexible and appropriate for survey datasets.

3. **Reduces Bias Compared to Listwise Deletion**  
   - removing missing cases could introduce bias. ML imputation minimizes this risk by estimating missing values based on observed data.

By using ML imputation, I ensure that my dataset remains as complete and unbiased as possible while preserving the integrity of the relationships between key variables.


```{r}
library(mice)   

# Perform multiple imputation (5 imputations by default)
imputed_data <- mice(subset_data, m = 1, method = "pmm", seed = 123)

```

```{r}
factor_vars <- c("year", "marital", "educ", "degree", "padeg", "madeg", "sex", 
                 "race", "reg16", "income", "rincome", "region", "partyid", 
                 "polviews", "relig", "happy", "class", "satfin", "finalter", "finrela")

subset_data_imputed <- subset_data_imputed %>%
  mutate(across(all_of(factor_vars), as.factor))

```






