---
title: "Regressions"
author: "Hugo van den Belt"
date: "2025-02-19"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

# Regressions

```{r}
# Packages
library(MASS)
library(dplyr)
```


## Function for formatted regression output
```{r}
summarize_polr <- function(model) {
  # Extract coefficients, standard errors, and t-values
  coefs <- coef(summary(model))
  
  # p-values two-tailed
  p_values <- (1 - pnorm(abs(coefs[, "t value"]))) * 2  
  
  # Function to add significance stars
  add_stars <- function(p) {
    if (p < 0.001) return("***")
    else if (p < 0.01) return("**")
    else if (p < 0.05) return("*")
    else if (p < 0.1) return(".")
    else return("")
  }
  
  # Apply the stars
  stars <- sapply(p_values, add_stars)
  
  # Round to 5 decimals 
  results_table <- cbind(
    round(coefs, 5),
    "p-value" = round(p_values, 5), 
    "Significance" = stars 
  )
  
  print(results_table, digits = 5)
  
  # Exponentiated coefficients 
  exp_coefs <- round(exp(coef(model)), 5)
  
  cat("\n**Exponentiated Coefficients (Odds Ratios):**\n")
  print(exp_coefs)
}

```


## Ordered Logit Models – Full Sample (`subset_final`)

### Republican President | Any Partisanship
```{r}
model_RepPres_FullSample <- polr(finalter ~ pres_republican + Democrat_any + Republican_any +
                        pres_republican:Democrat_any + pres_republican:Republican_any +
                        age + factor(sex) + factor(race) + factor(income) + factor(class) + factor(degree),
                        data = subset_final, Hess = TRUE)

summarize_polr(model_RepPres_FullSample)
```

### Democratic President | Any Partisanship
```{r}
model_DemPres_FullSample <- polr(finalter ~ pres_democrat + Democrat_any + Republican_any +
                        pres_democrat:Democrat_any + pres_democrat:Republican_any +
                        age + factor(sex) + factor(race) + factor(income) + factor(class) + factor(degree),
                        data = subset_final, Hess = TRUE)

summarize_polr(model_DemPres_FullSample)

```

### Republican President | Strong Partisanship
```{r}
model_RepPres_FullSample_Strong <- polr(finalter ~ pres_republican + Democrat_strong + Republican_strong +
                        pres_republican:Democrat_strong + pres_republican:Republican_strong +
                        age + factor(sex) + factor(race) + factor(income) + factor(class) + factor(degree),
                        data = subset_final, Hess = TRUE)

summarize_polr(model_RepPres_FullSample_Strong)

```

### Democratic President | Strong Partisanship
```{r}
model_DemPres_FullSample_Strong <- polr(finalter ~ pres_democrat + Democrat_strong + Republican_strong +
                        pres_democrat:Democrat_strong + pres_democrat:Republican_strong +
                        age + factor(sex) + factor(race) + factor(income) + factor(class) + factor(degree),
                        data = subset_final, Hess = TRUE)

summarize_polr(model_DemPres_FullSample_Strong)

```


## Ordered Logit Models – Post-2001 Sample (`subset_final_postbush`)

### Republican President | Any Partisanship
```{r}
model_RepPres_Post2001 <- polr(finalter ~ pres_republican + Democrat_any + Republican_any +
                        pres_republican:Democrat_any + pres_republican:Republican_any +
                        age + factor(sex) + factor(race) + factor(income) + factor(class) + factor(degree),
                        data = subset_final_postbush, Hess = TRUE)

summarize_polr(model_RepPres_Post2001)

```

### Democratic President | Any Partisanship
```{r}
model_DemPres_Post2001 <- polr(finalter ~ pres_democrat + Democrat_any + Republican_any +
                        pres_democrat:Democrat_any + pres_democrat:Republican_any +
                        age + factor(sex) + factor(race) + factor(income) + factor(class) + factor(degree),
                        data = subset_final_postbush, Hess = TRUE)

summarize_polr(model_DemPres_Post2001)

```

### Republican President | Strong Partisanship
```{r}
model_RepPres_Post2001_Strong <- polr(finalter ~ pres_republican + Democrat_strong + Republican_strong +
                        pres_republican:Democrat_strong + pres_republican:Republican_strong +
                        age + factor(sex) + factor(race) + factor(income) + factor(class) + factor(degree),
                        data = subset_final_postbush, Hess = TRUE)

summarize_polr(model_RepPres_Post2001_Strong)

```

### Democratic President | Strong Partisanship
```{r}
model_DemPres_Post2001_Strong <- polr(finalter ~ pres_democrat + Democrat_strong + Republican_strong +
                        pres_democrat:Democrat_strong + pres_democrat:Republican_strong +
                        age + factor(sex) + factor(race) + factor(income) + factor(class) + factor(degree),
                        data = subset_final_postbush, Hess = TRUE)

summarize_polr(model_DemPres_Post2001_Strong)

```


## Predicted Probabilites

```{r}
# This function will get me modal/mean values of covariates from a to-be specified df
get_representative_values <- function(data) {
  list(
    modal_sex = as.character(names(which.max(table(data$sex)))),
    modal_race = as.character(names(which.max(table(data$race)))),
    modal_income = as.character(names(which.max(table(data$income)))),
    modal_class = as.character(names(which.max(table(data$class)))),
    modal_degree = as.character(names(which.max(table(data$degree)))),
    mean_age = mean(data$age, na.rm = TRUE)
  )
}

# Function to create profiles based on partisan combinations
create_partisan_profiles <- function(rep_values, partisan_type = "any") {
  # modal/mean values
  base_profile <- list(
    sex = rep_values$modal_sex,
    race = rep_values$modal_race,
    income = rep_values$modal_income,
    class = rep_values$modal_class,
    degree = rep_values$modal_degree,
    age = rep_values$mean_age
  )
  
  # Different partisan profiles based on partisan_type
  if (partisan_type == "any") {
    profiles <- list(
      rep_rep = c(base_profile, list(pres_republican = 1, Democrat_any = 0, Republican_any = 1)),
      rep_dem = c(base_profile, list(pres_republican = 0, Democrat_any = 0, Republican_any = 1)),
      dem_rep = c(base_profile, list(pres_republican = 1, Democrat_any = 1, Republican_any = 0)),
      dem_dem = c(base_profile, list(pres_republican = 0, Democrat_any = 1, Republican_any = 0))
    )
  } else if (partisan_type == "strong") {
    profiles <- list(
      rep_rep = c(base_profile, list(pres_republican = 1, Democrat_strong = 0, Republican_strong = 1)),
      rep_dem = c(base_profile, list(pres_republican = 0, Democrat_strong = 0, Republican_strong = 1)),
      dem_rep = c(base_profile, list(pres_republican = 1, Democrat_strong = 1, Republican_strong = 0)),
      dem_dem = c(base_profile, list(pres_republican = 0, Democrat_strong = 1, Republican_strong = 0))
    )
  }
  
  # Each profile has to convert to a data frame
  profiles <- lapply(profiles, function(p) {
    df <- data.frame(p)
    # Ensure factors are properly set for prediction
    for (col in c("sex", "race", "income", "class", "degree")) {
      if (col %in% names(df)) {
        df[[col]] <- as.factor(df[[col]])
      }
    }
    return(df)
  })
  
  return(profiles)
}

# Fredicted probabilities for a given model, and profiles
calculate_predicted_probs <- function(model, profiles) {
  # Level names for finalter
  level_names <- levels(model$model$finalter)
  
  # Predicted probabilities for each profile
  pred_probs <- list()
  for (name in names(profiles)) {
    pred_probs[[name]] <- predict(model, newdata = profiles[[name]], type = "probs")
  }
  
  # results table
  result_table <- data.frame(
    partisan_pres = c("Republican under Republican President", 
                      "Republican under Democratic President",
                      "Democrat under Republican President",
                      "Democrat under Democratic President")
  )
  
  # Add the probability columns with correct names
  for (i in 1:length(level_names)) {
    col_name <- level_names[i]
    result_table[[col_name]] <- c(pred_probs$rep_rep[i], 
                                 pred_probs$rep_dem[i], 
                                 pred_probs$dem_rep[i], 
                                 pred_probs$dem_dem[i])
  }
  
  # Make into percentages
  result_table[, -1] <- round(result_table[, -1] * 100, 1)
  
  return(result_table)
}

# Main function (all predicted probabilities for a model)
get_predicted_probabilities <- function(model, data, partisan_type = "any") {
  rep_values <- get_representative_values(data)
  profiles <- create_partisan_profiles(rep_values, partisan_type)
  prob_table <- calculate_predicted_probs(model, profiles)
  
  return(prob_table)
}

# Function to run and store predicted probabilities for all model combinations
get_all_predicted_probabilities <- function(subset_final, subset_final_postbush) {
  results <- list()
  
  # Models for full sample with any partisanship
  model_RepPres_FullSample <- polr(finalter ~ pres_republican + Democrat_any + Republican_any +
                        pres_republican:Democrat_any + pres_republican:Republican_any +
                        age + factor(sex) + factor(race) + factor(income) + factor(class) + factor(degree),
                        data = subset_final, Hess = TRUE)
  
  results$full_any <- get_predicted_probabilities(model_RepPres_FullSample, subset_final, "any")
  
  # Models for full sample with strong partisanship
  model_RepPres_FullSample_Strong <- polr(finalter ~ pres_republican + Democrat_strong + Republican_strong +
                        pres_republican:Democrat_strong + pres_republican:Republican_strong +
                        age + factor(sex) + factor(race) + factor(income) + factor(class) + factor(degree),
                        data = subset_final, Hess = TRUE)
  
  results$full_strong <- get_predicted_probabilities(model_RepPres_FullSample_Strong, subset_final, "strong")
  
  # Models for post-2001 sample with any partisanship
  model_RepPres_Post2001 <- polr(finalter ~ pres_republican + Democrat_any + Republican_any +
                        pres_republican:Democrat_any + pres_republican:Republican_any +
                        age + factor(sex) + factor(race) + factor(income) + factor(class) + factor(degree),
                        data = subset_final_postbush, Hess = TRUE)
  
  results$post2001_any <- get_predicted_probabilities(model_RepPres_Post2001, subset_final_postbush, "any")
  
  # Models for post-2001 sample with strong partisanship
  model_RepPres_Post2001_Strong <- polr(finalter ~ pres_republican + Democrat_strong + Republican_strong +
                        pres_republican:Democrat_strong + pres_republican:Republican_strong +
                        age + factor(sex) + factor(race) + factor(income) + factor(class) + factor(degree),
                        data = subset_final_postbush, Hess = TRUE)
  
  results$post2001_strong <- get_predicted_probabilities(model_RepPres_Post2001_Strong, subset_final_postbush, "strong")
  
  return(results)
}

```


### Results

```{r}
# Run all the calculations and store results
all_results <- get_all_predicted_probabilities(subset_final, subset_final_postbush)

print(all_results)
```

