---
title: "Robustness of Regressions"
author: "Hugo van den Belt"
date: "2025-02-19"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

# Robustness of Regressions

In this file, I run extra regressions to check the robustness of findings.

```{r}
library(MASS)

summarize_polr <- function(model) {
  # Extract coefficients, standard errors, and t-values
  coefs <- coef(summary(model))
  
  # p-values two-tailed
  p_values <- (1 - pnorm(abs(coefs[, "t value"]))) * 2  
  
  # Function to add significance stars
  add_stars <- function(p) {
    if (p < 0.001) return("***")
    else if (p < 0.01) return("**")
    else if (p < 0.05) return("*")
    else if (p < 0.1) return(".")
    else return("")
  }
  
  # Apply the stars
  stars <- sapply(p_values, add_stars)
  
  # Round to 5 decimals 
  results_table <- cbind(
    round(coefs, 5),
    "p-value" = round(p_values, 5), 
    "Significance" = stars 
  )
  
  print(results_table, digits = 5)
  
  # Exponentiated coefficients 
  exp_coefs <- round(exp(coef(model)), 5)
  
  cat("\n**Exponentiated Coefficients (Odds Ratios):**\n")
  print(exp_coefs)
}

```


## Regressions with omitted years (1977, 1993, 2021)

### Republican President | Any Partisanship
```{r}
model_RepPres_FullSample <- polr(finalter ~ pres_republican + Democrat_any + Republican_any +
                        pres_republican:Democrat_any + pres_republican:Republican_any +
                        age + factor(sex) + factor(race) + factor(income) + factor(class) + factor(degree),
                        data = subset_data_clean, Hess = TRUE)

summarize_polr(model_RepPres_FullSample)
```


### Republican President | Strong Partisanship
```{r}
model_RepPres_FullSample_Strong <- polr(finalter ~ pres_republican + Democrat_strong + Republican_strong +
                        pres_republican:Democrat_strong + pres_republican:Republican_strong +
                        age + factor(sex) + factor(race) + factor(income) + factor(class) + factor(degree),
                        data = subset_data_clean, Hess = TRUE)

summarize_polr(model_RepPres_FullSample_Strong)

```



## Regressions with time covariate

### Republican President | Any Partisanship
```{r}
model_RepPres_FullSample <- polr(finalter ~ pres_republican + Democrat_any + Republican_any +
                        pres_republican:Democrat_any + pres_republican:Republican_any +
                        age + factor(sex) + factor(race) + factor(income) + factor(class) + factor(degree) + factor(year),
                        data = subset_final, Hess = TRUE)

summarize_polr(model_RepPres_FullSample)
```



### Republican President | Strong Partisanship
```{r}
model_RepPres_FullSample_Strong <- polr(finalter ~ pres_republican + Democrat_strong + Republican_strong +
                        pres_republican:Democrat_strong + pres_republican:Republican_strong +
                        age + factor(sex) + factor(race) + factor(income) + factor(class) + factor(degree) + factor(year),
                        data = subset_final, Hess = TRUE)

summarize_polr(model_RepPres_FullSample_Strong)

```

### Republican President | Any Partisanship | Post-Bush
```{r}
model_RepPres_Post2001 <- polr(finalter ~ pres_republican + Democrat_any + Republican_any +
                        pres_republican:Democrat_any + pres_republican:Republican_any +
                        age + factor(sex) + factor(race) + factor(income) + factor(class) + factor(degree) + factor(year),
                        data = subset_final_postbush, Hess = TRUE)

summarize_polr(model_RepPres_Post2001)

```


### Republican President | Strong Partisanship | Post-Bush
```{r}
model_RepPres_Post2001_Strong <- polr(finalter ~ pres_republican + Democrat_strong + Republican_strong +
                        pres_republican:Democrat_strong + pres_republican:Republican_strong +
                        age + factor(sex) + factor(race) + factor(income) + factor(class) + factor(degree) + factor(year),
                        data = subset_final_postbush, Hess = TRUE)

summarize_polr(model_RepPres_Post2001_Strong)

```















